Eddy covariance towers as sentinels of abnormal radioactive material releases

### 2024-11-12 manuscript/manuscript.tex
stash edits

For both the IRE and Fukushima case studiesy, we gathered measurements from eddy covariance platforms (i.e., “flux towers”) spanning multiple years. For the IRE case study, we targeted towers located in Belgium, which are part of the ICOS (Integrated Carbon Observation System) network available through the European Fluxes Database (http://www.europe-fluxdata.eu/). Three sites in proximity to the known release location had sufficient long-term data for our anticipated data mining efforts (BE-Lon, BE-Bra, and BE-Vie sites, Figure 1A). Each of the three flux tower sites are located within a crop ecosystem type and had almost 10 years of data available from 2004 through 2013 recorded at 30 min intervals. While BE-Lon had a near-continuous data record during this period, BE-Bra had a long period of missing data in 2005, and both BE-Bra and BE-Vie had a long period of missing data in 2009. The distance between IRE and the flux tower sites are approximately 19 km, 95 km, and 105 km for the BE-Lon, BE-Bra, and BE-Vie, respectively.

@@

For the IRE case study, among all possible pairwise combinations of dependent and independent variables (n=51, Table S1), we found that only the $CO_2$ versus air temperature comparison had any measure of predictability from the model and a strong event effect size (Table 1). Note that values in Table S1 are computed relative to the entirety of the data record for variable pair screening purposes, while the values in Table 1 are computed over the known abnormal event period for evaluating event identifiability. Although other variable pairs such as sensible heat flux and net radiation had a stronger effect size for the BE-Lon site, they had a weak linear relationship across the other sites. When we focus on the period of the abnormal release event in particular rather than the overall period of record, we found that $CO_2$ versus air temperature relationship at each of the three focal towers had similar $R^2$ values and effect sizes (Table 1). Figure 2 provides a visual example of the effect size framing in our analysis. Note how the apparent interaction effect between time periods around the abnormal release event in Figure 2A (i.e., different slopes for before, during, after) translates to a relatively high effect size (Figure 2C) whereas in Figure 2B at an arbitrary time point, there is no apparent interaction effect. This corresponds to a low effect size in Figure 2D. Note also how during and after the known abnormal release, the slopes of the relationships between $CO_2$ and air temperature become less negative than before the event, with lower than expected $CO_2$ at low air temperature (Figure 2A). A similar pattern is apparent for the BE-Bra and BE-Vie sites (Figure S1, S2).

@@

To increase our confidence that event detections were not dependent on the specific hyperparameters settings in our base-case setup, we ran an exhaustive cross-validation set of hyperparameter experiments. For this effort, we tested every possible combination of values for wind tolerance, number of days, and the event quantile effect. Across reasonable values of specific hyperparameters, we found that their magnitudes had little effect on the rate at which events were flagged as potential abnormal releases as even in the most extreme cases, we did not see detection rates exceeding 1\%. We did, however, observe that the event detection was related to the window length and event quantile parameters as shown by the slope of the line in Figure 4. Although increasing values of the event quantile parameter above the limit of our tests ($>$ 0.9) would hypothetically increase event detection rates, we do not show this due to the noisiness of the extreme quantiles of effect size distributions. Additionally, we do not show tests setting this parameter below the median because it would lead to  weak sensitivity to abnormal events. In a similar fashion as the event quantile parameter, increasing the magnitude of the window length parameter would also likely increase the event detection rate but here too we do not show tests setting this parameter beyond 14 days given the results of prior atmospheric modeling efforts showing limited deposition at such long time scales \citep[e.g.][]{meszarosPredictabilityDispersionFukushimaderived2016}.

@@

More generally, we show that flux tower records have value for this type of data mining despite the apparent noisiness of the data \citep[e.g., ][]{fratiniEddyCovarianceFlux2018}. We attribute the sensitivity of our approach to the fact that we fit successive interaction models to limited sections of the data record, and we leveraged bivariate relationships to constrain noisiness in individual data records. Our approach was aided by the fact that this type of scale-free data mining is largely unaffected by systematic errors due to uncertainty in calibration standards and/or low resolution in specific sensor packages deployed on the towers. This is because although such systematic errors may bias the magnitude of overall flux values, which would be a problem for typical uses of flux towers that deal directly in the units of the measured data, they do not reduce the overall confidence in individual values \citep{langfordEddycovarianceDataLow2015}, which for our purposes might represent an abnormal release signal. The ability of our approach to identify a suitable release signature does likely depend on the quality and completeness of the underlying data.

>>>>

For both the IRE and Fukushima case studies, we gathered measurements from eddy covariance platforms (i.e., “flux towers”) spanning multiple years. For the IRE case study, we targeted towers located in Belgium, which are part of the ICOS (Integrated Carbon Observation System) network available through the European Fluxes Database (http://www.europe-fluxdata.eu/). Three sites in proximity to the known release location had sufficient long-term data for our anticipated data mining efforts (BE-Lon, BE-Bra, and BE-Vie sites, Figure 1A). Each of the three flux tower sites are located within a crop ecosystem type and had almost 10 years of data available from 2004 through 2013 recorded at 30 min intervals. While BE-Lon had a near-continuous data record during this period, BE-Bra had a long period of missing data in 2005, and both BE-Bra and BE-Vie had a long period of missing data in 2009. The distance between IRE and the flux tower sites are approximately 19 km, 95 km, and 105 km for the BE-Lon, BE-Bra, and BE-Vie, respectively.

@@

For the IRE case study, among all possible pairwise combinations of dependent and independent variables (n=51, Table S1), we found that only the $CO_2$ versus air temperature comparison had any measure of predictability from the model and a strong event effect size (Table 1). Note that values in Table S1 are computed relative to the entirety of the data record for variable pair screening purposes, while the values in Table 1 are computed over the known abnormal event period for event identifiability purposes. Although other variable pairs such as sensible heat flux and net radiation had a stronger effect size for the BE-Lon site, they had a weak linear relationship across the other sites. When we focus on the period of the abnormal release event in particular rather than the overall period of record, we found that $CO_2$ versus air temperature relationship at each of the three focal towers had similar $R^2$ values and effect sizes (Table 1). Figure 2 provides a visual example of the effect size framing in our analysis. Note how the apparent interaction effect between time periods around the abnormal release event in Figure 2A (i.e., different slopes for before, during, after) translates to a relatively high effect size (Figure 2C) whereas in Figure 2B at an arbitrary time point, there is no apparent interaction effect. This corresponds to a low effect size in Figure 2D. Note also how during and after the known abnormal release, the slopes of the relationships between $CO_2$ and air temperature become less negative than before the event, with lower than expected $CO_2$ at low air temperature (Figure 2A). A similar pattern is apparent for the BE-Bra and BE-Vie sites (Figure S1, S2).

@@

To increase our confidence that event detections were not dependent on the specific hyperparameters settings in our base-case setup, we ran an exhaustive cross-validation set of hyperparameter experiments. For this effort, we tested every possible combination of values for wind tolerance, number of days, and the event quantile effect. Across reasonable values of specific hyperparameters, we found that their magnitudes had little effect on the rate at which events were flagged as potential abnormal releases as even in the most extreme cases, we did not see detection rates exceeding 1\%. We did, however, observe that the event detection was related to the window length and event quantile parameters as shown by the slope of the line in Figure 4. Although increasing values of the event quantile parameter above the limit of our tests ($>$ 0.9) would hypothetically increase event detection rates, we do not show this due to the noisiness of the extreme quantiles of effect size distributions. Additionally, we do not show tests setting this parameter below the median because it would lead to weak sensitivity to abnormal event detection. In a similar fashion as the event quantile parameter, increasing the magnitude of the window length parameter would also likely increase the event detection rate but here too we do not show tests setting this parameter beyond 14 days given the results of prior atmospheric modeling efforts showing limited deposition at such long time scales \citep[e.g.][]{meszarosPredictabilityDispersionFukushimaderived2016}.

@@

More generally, we show that flux tower records have value for this type of data mining despite the apparent noisiness of the data \citep[e.g., ][]{fratiniEddyCovarianceFlux2018}. We attribute the sensitivity of our approach to the fact that we fit successive interaction models to limited sections of the data record, and we leveraged bivariate relationships to constrain noisiness in individual data records. This strategy of using only limited sections of the data is particularly important because it can avoid the situation where a well-defined linear relationship becomes non-linear and obscured at longer timescales. Despite the apparent success of our approach, it is not a perfect solution because some non-linearity remains. Beyond the noise inherent to most environmental data, this is likely one of the principal causes of the relatively modest $R^2$ values we observed in Table 1. These are not unusally low however as $R^2$ values between 0.35 and 0.7 are typical for environmental data \citep{murrayMethodsQuantifyVariable2009}. A fruitful area of future research would be exploring physical justifications for linearizing these relationships, which would likely lead to more precise signal recovery and event detection. Our approach was further aided by the fact that this type of scale-free data mining is largely unaffected by systematic errors due to uncertainty in calibration standards and/or low resolution in specific sensor packages deployed on the towers. This is because although such systematic errors may bias the magnitude of overall flux values, which would be a problem for typical uses of flux towers that deal directly in the units of the measured data, they do not reduce the overall confidence in individual values \citep{langfordEddycovarianceDataLow2015}, which for our purposes might represent an abnormal release signal. As a result, the ability of our approach to identify a suitable release signature does depend on the quality and completeness of the underlying data.
